---
title: "MosaicPICASSO Tutorial"
author: "Edward Schrom"
date: "2024-01-10"
output: html_document
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{MosaicPICASSO Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

MosaicPICASSO infers and removes artefacts of autofluorescence and spillover from multiplex images of biological tissues. There is no need for extra control images of autofluorescence only, single color staining, etc. 

In brief, the PICASSO algorithm assumes that the artefactual contributions of autofluorescence and spillover are detectable as shared signal across channels. Thus, for every ordered pair of channels (i,j), a scaling factor "a" is chosen to subtract a*channel i from channel j. The scaling factor is chosen to minimize the normalized mutual information between the two images. This process is iterated over all pairs of channels to correct the entire image. This algorithm was published in [@seo2022].

Alone, the PICASSO algorithm sometimes subtracts too much true signal, particularly when channels genuinely overlap spatially. The Mosaic addition to the algorithm solves this by first breaking the image into many tiles. For each ordered pair of channels, the structural similarity index measure (SSIM) is calculated for corresponding tiles. Only tiles below a threshold quantile of SSIM are used to calculate the scaling factor for the pair of channels. Thus, scaling factors are calculated based only on portions of the image without genuine overlap between channels. The Mosaic addition was published in [@cang2023].

This R package adds a few minor features to the complete MosaicPICASSO algorithm:

- Absolute value of correlation, rather than normalized mutual information, can be chosen as the metric to minimize.
- The pixels used to calculate the minimization metric can be subsampled to improve computational speed.
- The corrected image can be contrast-enhanced to match the saturation percentage of the original image.

# Usage

MosaicPICASSO operates only on 8-bit grayscale multichannel .tif files. Using MosaicPICASSO requires only a single command, of the following format:

```{r eval=FALSE}
MosaicPICASSO::mosaic_picasso("path/to/original/image.tif", method="MI", q_thr=0.5, tile_size=50, stride=50, pix_ss="all", enhance=T, cores=NULL)

```

No output is produced in R. Instead, two files are saved to the working directory:

- A corrected version of the image, saved as a .tif with the original file name plus "_NEW"
- A copy of the correction matrix, saved as a .csv with the original file name plus "_MAT"

# Parameters

method: “MI” or “Cor” [default “MI”]. Metric to minimize when choosing the scaling factor.

- "MI" can produce more accurate output, but with much slower computational times
- "Cor" can produce less accurate output, but with much faster computational times

q_thr: number from 0 to 1 [default 0.5]. Fraction of tiles to use with lowest SSIM.

- Use a lower value (e.g. 0.1) if markers overlap substantially, so that only a small part of image with the least true overlap is used to calculate scaling factors
- Use a higher value (e.g. 0.9) if markers do not overlap substantially, so that most of the image except where true overlap occurs is used to calculate scaling factors

tile_size: number of pixels [default 50]. Size of square tiles.

- Choose a value that captures 1-10 cells, based on the image resolution

stride: number of pixels [default 50]. Space between top left corners of square tiles.

- If stride > tile_size, there will be gaps between the tiles
- If stride < tile_size, the tiles will overlap

pix_ss: number of pixels [default “all”]. Number of pixels to subsample.

- For large images (>> 10,000,000 pixels, roughly), set to 100,000 or 1,000,000 to speed up calculations.


cycles: whole number [default 1]. Number of rounds of correction to perform.

- The output image can be run through mosaic-PICASSO again (and again, etc.)
- This is rarely warranted and should usually be set to 1

enhance: true or false [default T]. Whether to enhance contrast of new image.

- Mosaic-PICASSO can boost signal:noise substantially, but its channel subtraction can only makes images dimmer at large
- Enhancing contrast saturates the same percentage of pixels as in the original image, for each channel separately

cores: whole number [default NULL]. Number of CPU cores for parallelization.

- Parallel computations make Mosaic-PICASSO faster; by default it will use all the cores you have, minus 1
- You can specify an exact number of cores to bypass the default

# References